---
title: "DataCleaning"
author: "Jacob Johns"
date: "2023-11-03"
output: html_document
---

# Data cleaning

## Initialize Document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
```

## Read In Data

```{r}
traffic = read.csv("Road_Traffic_Counts_in_Colorado_2019_20231027.csv")
```

## Initial Feature Extraction

Filter out the 29,373 observations of road outside of city of Denver

```{r}
traffic = subset(traffic, traffic$FIPS=="Denver")
```

Filter out the 144 observations of Freeways

```{r}
traffic = subset(traffic,traffic$FUNCCLASSI !="2  Principal Arterial - Fwys and Expwys")
```

Filter out features which are same for all Denver roads

```{r}
traffic = subset(traffic, select = -c(the_geom,FIPS,FIPSCOUNTY,HPMSID,URBAN,POPULATION,SPECIALSYS,GOVLEVEL,AADTYR,FORESTROUT,FundID))
```

Filter out features which don't distinguish enough

```{r}
traffic = subset(traffic, select = -c(GISID,ADMINCLASS,SURF_TYPE,OPERATION,NHSDESIG,JURSPLIT,LRSROUTE,BUILTYR))
```

Filter out features which we're not interested in

```{r}
traffic = subset(traffic, select = -c(ROUTE,FROM_DESCR,TO_DESCR,SEGMID,IRI,COUNTSTATI))
```

## Factorize Qualitative Features

We want to be able to test if qualitative features (like direction of the street) affects traffic. So we must consider them as factors, (features that have levels) to analyze their correlation.

```{r}
traffic$SEGMDIR = as.factor(traffic$SEGMDIR)
traffic$FUNCCLASSI = as.factor(traffic$FUNCCLASSI)
traffic$SURFNAME = as.factor(traffic$SURFNAME)
```

## Summarize

```{r}
summary(traffic)
```

```{r}
head(traffic)
```

## Remove NA

There aren't any NA values, so there is nothing to worry about in this aspect.

```{r}
traffic = na.omit(traffic)
```

## Test/Train Split

```{r}
set.seed(1)
traffic$id = 1:nrow(traffic)
trafficTrain <- traffic %>% sample_frac(0.80)
trafficTest  <- anti_join(traffic, trafficTrain, by = 'id')
```

## Export

```{r}
write.csv(trafficTest , file = "DenTraffic2019Test.csv")
write.csv(trafficTrain , file = "DenTraffic2019Train.csv")
```
