---
title: "LinearProgrammingPart"
author: "Jacob Johns"
date: "2023-11-21"
output: html_document
---

# Linear Programming

## Initialize Document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read In Data

```{r}
traffic = read.csv("Road_Traffic_Counts_in_Colorado_2019_20231027.csv")
```

## Add Network Flow Constraints

```{r}
traffic$CAPACITY = traffic$AADT*traffic$SEG_LENGTH/24/25
traffic$COST = traffic$SEG_LENGTH
traffic = subset(traffic, select = c(SEGMID,COST,FUNCCLASSI,SURFWD,THRULNQTY,CAPACITY,AADT))
```

## Subset Surveyed Roads

```{r}
lincoln = subset(traffic,rownames(traffic)%in%c(4142,3098,845,2049,2286))
grant = subset(traffic,rownames(traffic)%in%c(7957,9186,9831,5769,9063))
wash = subset(traffic,rownames(traffic)%in%c(6443,7679,9745,9853,7791))
clark = subset(traffic,rownames(traffic)%in%c(8314,6550,5490,8545,9848))
corona = subset(traffic,rownames(traffic)%in%c(7786,5093,6483,6077,9137))
logan = subset(traffic,rownames(traffic)%in%c(6411,8116,6525,7290,5420))
thirteenth = subset(traffic,rownames(traffic)%in%c(5740,5506,8160,6734,7802,7567,9864))
eigth = subset(traffic,rownames(traffic)%in%c(6899,9973,9572,7914,6270,5846,6440,9337))
eleventh = subset(traffic,rownames(traffic)%in%c(25945,26021,25977,25361,25392,25794,25206,25598,25692,26047))
```

## Read In Model

```{r}
model = readRDS("model3Limited")
```

## Helper Function for Predicting

```{r}
generate_df <- function(SEG_LENGTH,SURFWD,model,THRULNQTY=2,FUNCCLASSI ="5  Major Collector"){
  street = data.frame(FUNCCLASSI,SURFWD,THRULNQTY,SEG_LENGTH)
  street$AADT = predict(model, type = 'response', newdata = street)
  street$CAPACITY = street$AADT*street$SEG_LENGTH/24/25
  street$COST = street$SEG_LENGTH
  street = subset(street, select = c(COST,THRULNQTY,CAPACITY,AADT))
  return(street)
}
```

## Predict Current Non-Surveyed Roads

```{r}
SEG_LENGTH = c(0.14,0.07,0.19,0.06,0.19)
SURFWD = c(32,34,32,36,36)
predNinethE = generate_df(SEG_LENGTH,SURFWD,model)
predNinethW = generate_df(rev(SEG_LENGTH),rev(SURFWD),model)

SURFWD = c(32,36,36,36,36)
predTenthE = generate_df(SEG_LENGTH,SURFWD,model)
predTenthW = generate_df(rev(SEG_LENGTH),rev(SURFWD),model)

SURFWD = c(34,24,36,36,36)
predTwelfthE = generate_df(SEG_LENGTH,SURFWD,model)
predTwelfthW = generate_df(rev(SEG_LENGTH),rev(SURFWD),model)
```

## Predict One-way Conversion

If we were to make 9th and 12th east bound one-ways and 10th a west bound one-way, we would expect the following data, assuming we now have 3 through lanes.

```{r}
SEG_LENGTH = c(0.14,0.07,0.19,0.06,0.19)
SURFWD = c(32,34,32,36,36)
predNinethEOne = generate_df(SEG_LENGTH,SURFWD,model,THRULNQTY = 3)

SURFWD = c(32,36,36,36,36)
predTenthWOne = generate_df(rev(SEG_LENGTH),rev(SURFWD),model,THRULNQTY = 3)

SURFWD = c(34,24,36,36,36)
predTwelfthEOne = generate_df(SEG_LENGTH,SURFWD,model,THRULNQTY = 3)
```

## Predict Two-way Conversion

If we were to make Washington, Clarkson, and Corona (because they primarly have housing units on them) two-ways, this is what we'd expect, assuming

```{r}
SEG_LENGTH = c(0.10,0.11,0.11,0.10,0.12)
SURFWD = c(32,32,32,32,32)
predWashNTwo = generate_df(SEG_LENGTH,SURFWD,model,THRULNQTY = 1)
predWashSTwo = generate_df(rev(SEG_LENGTH),rev(SURFWD),model,THRULNQTY = 1)


predClarkNTwo = generate_df(SEG_LENGTH,SURFWD,model,THRULNQTY = 1)
predClarkSTwo = generate_df(rev(SEG_LENGTH),rev(SURFWD),model,THRULNQTY = 1)


predCorNTwo = generate_df(SEG_LENGTH,SURFWD,model,THRULNQTY = 1)
predCorSTwo = generate_df(rev(SEG_LENGTH),rev(SURFWD),model,THRULNQTY = 1)
```
