---
title: "Exploration"
author: "Jacob Johns"
date: "2023-11-03"
output: html_document
---

# Data Exploration

## Initialize Document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(reshape2)
```

## Read In Data

```{r}
traffic = read.csv("DenTraffic2019Train.csv")
traffic$SEGMDIR = as.factor(traffic$SEGMDIR)
traffic$FUNCCLASSI = as.factor(traffic$FUNCCLASSI)
traffic$SURFNAME = as.factor(traffic$SURFNAME)
```

## Feature Correlation

We will run a correlation on each of our quantitative features to ensure that we avoid the issue of multiple collinearity. We choose an alpha level of 0.70 so any features that share a correlation higher than that, one of them will be chosen to be omitted.

```{r}
numericData = subset(traffic, select = -c(X,ROUTENAME,SEGMDIR,FUNCCLASSI,SURFNAME,id))
cormat = round(cor(numericData),2)
cormat <- melt(cormat)
heatMap = ggplot(data = cormat, aes(x=Var1,y=Var2,fill=value))
heatMap = heatMap + geom_tile() 
heatMap = heatMap + scale_fill_gradient2(low = "blue", high = "red", mid = "white",
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation")
heatMap = heatMap + theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 10, hjust = 1))
heatMap = heatMap + geom_text(aes(Var2, Var1, label = value), color = "black", size = 3)
heatMap
```

Here we see that $\verb|RUNLENGTH_|$ and $\verb|RUNLENGTH1|$ are actually the same variable. Accordingly, we'll remove one of them. It appears that there are no other correlations to be concerned about, but we will verify that later in the modeling section.

## Data Quality

```{r}
barplot(table(traffic$AADTDERIV))
```

This is shows how many years since the $\verb|AADT|$ data was collected, except for the values of 4, which are defined by the data set to be estimated. We believe using estimated/calculated values to predict data will lead the model to fit the estimation the dataset owners used. Accordingly, we will remove those observations.

```{r}
traffic = subset(traffic,traffic$AADTDERIV!=4)
```

## Summarization

```{r}
summary(traffic)
```

From our summaries, we can see that there isn't really enough support for using $\verb|SURFNAME|$ as a predictor variable

## Feature Selection

```{r}
traffic = subset(traffic, select = -c(X,RUNLENGTH_,id,SURFNAME))
```

## Export

```{r}
write.csv(traffic , file = "DenTraffic2019ToModel.csv")
```
